---
title: "BMI Health App Presentation"
author: "Ruth Lee"
date: "April 1, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Heath Indicator using Body Mass Index (BMI)

**Body Mass Index (BMI)** is a simple health indicator using the height and weight of a person. The formula is BMI = a person's weight (kg) divided by the square of his height (m).

This simple app calculates your BMI and the amount of weight to gain (if you are underweight) or to lose (if you are overweight or obese).
        
Note that BMI is not appropriate for use as a health indicator for pregnant women, the elderly, young children and muscle-builders.

## App Features

- Computes your BMI
- Tells you whether you have a normal body weight
- Computes your ideal weight based on your height
- Computes the amount of weight to gain or lose to reach normal body weight (ideal weight) range
- Shows where your weight is vis-a-vis the ideal weight


## R Code for the User Interface at ui.R


```{r code1, echo=TRUE,results='hide',message=FALSE}
library(shiny)
shinyUI(fluidPage(
sidebarPanel(
sliderInput('wt','Enter Your Weight (kg)',value=50,min= 20, max=200, step=0.05),
sliderInput('ht','Enter Your Height (cm)',value=150,min= 40, max=300, step=0.05),
submitButton("Submit")),
mainPanel(
  h2('Documentation on the Use of this App'), ## Details not displayed
  h2('Your BMI'), verbatimTextOutput("bmi"), verbatimTextOutput("text"),
  h2('Your Ideal Weight based on Your Height'),
  h5('Minimum (kg)'), verbatimTextOutput('idealmin'),
  h5('Maximum (kg)'), verbatimTextOutput('idealmax'),
  plotOutput("plot1"))))

```

## Output: Comparison Chart   
Pseudo Inputs: Weight = 40kg, Height = 160cm

```{r pseudo,echo=FALSE,comment="",fig.height = 4}
##pseudo weight and height figures
wt=40
ht=160
wtmin<-18.5*(ht/100)^2-wt
    wtmax<-24.9*(ht/100)^2-wt
    bmi<-wt/(ht/100)^2
    sprintf("Your BMI is %.2f.You are underweight.",bmi)
    sprintf("Please gain %.2f to %.2f kg to reach a normal body weight.", wtmin, wtmax)
    
    counts<-c(wt,18.5*(ht/100)^2, 24.9*(ht/100)^2)
    x = c("Your Weight", "Ideal Weight (Min)", "Ideal Weight (Max)")
   barplot(counts,horiz = FALSE,names.arg = x,col = c("red","blue","blue"),main="Your Weight vs Ideal Weight based on Your Height")
   abline(h=wt,col="red",lwd=2,lty=2)
```


```{r code, echo = FALSE}
library(shiny)
BMI<- function(wt,ht) wt/(ht/100)^2

shinyServer(function(input, output) {
  output$bmi<- renderPrint({BMI(input$wt,input$ht)})
  tt <- reactive({
  if(BMI(input$wt,input$ht)<18.5){
    wtmin<-18.5*(input$ht/100)^2-input$wt
    wtmax<-24.9*(input$ht/100)^2-input$wt
    text<-sprintf("You are underweight. Please gain %.2f to %.2f kg to reach a normal body weight.", wtmin, wtmax)
    }else if(BMI(input$wt,input$ht)<25) {
      text<-"You have a normal body weight. Continue to maintain this weight."
    }else if(BMI(input$wt,input$ht)<30){
      wtmin<-input$wt-18.5*(input$ht/100)^2
      wtmax<-input$wt-24.9*(input$ht/100)^2
      text<-sprintf("You are overweight. Please try to lose %.2f to %.2f kg to reach a normal body weight.", wtmax,wtmin)
    }else if(BMI(input$wt,input$ht)>30){
      wtmin<-input$wt- 18.5*(input$ht/100)^2
      wtmax<-input$wt- 24.9*(input$ht/100)^2
      text<-sprintf("You are obese. You should try to lose %.2f to %.2f kg to reach a normal body weight.", wtmax,wtmin)}
      })
    observe({
    input$wt
    input$ht})
  
  output$text<-renderText(tt())
    output$idealmin<-renderText({
    18.5*(input$ht/100)^2})
     
  output$idealmax<-renderText({
    24.8*(input$ht/100)^2})
  output$plot1 <- renderPlot({
    counts<-c(input$wt,18.5*(input$ht/100)^2, 24.9*(input$ht/100)^2)
    x = c("Your Weight", "Ideal Weight (Min)", "Ideal Weight (Max)")
   barplot(counts,horiz = FALSE,names.arg = x,col = c("red","blue","blue"),main="Your Weight vs Ideal Weight based on Your Height")
   abline(h=input$wt,col="red",lwd=2,lty=2)  })
  })
```